---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(message=FALSE, comment="#>")
devtools::load_all(".")
```

# shine

[![](https://img.shields.io/badge/platforms-linux%20%7C%20osx%20-2a89a1.svg)]()
[![](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![](https://img.shields.io/github/last-commit/montilab/shine.svg)](https://github.com/montilab/shine/commits/master)

**S**tructure Learning for **Hi**erarchical **Ne**tworks  
*A package to aid in structure learning for hierarchical biological regulatory networks*

## Documentation

Please visit https://montilab.github.io/shine/ for comprehensive documentation.

## Requirements

We suggest R 3.6.0 but R (>= 3.5.0) is required to install directly from Github. For workflows, you will need Python (>= 2.7.0) and dependencies for [Nextflow](https://www.nextflow.io/). Nextflow can be used on any POSIX compatible system (Linux, OS X, etc) and requires BASH and Java 8 (or higher) to be installed. Alternatively, check out usage with Docker.

## Installation

Install the development version of the package from Github.

```{r, eval=FALSE}
devtools::install_github("montilab/shine")
```

```{r, eval=FALSE}
library(shine)
```

## Quick Example

```{r}
data(toy)
```

```
    ABC
    / \
   AB  \ 
  /  \  \
 A    B  C 
```

```{r}
dim(toy)
table(toy$subtype)
```

### Variable Selection
```{r}
# Filter out non-varying genes
genes.filtered <- keep.var(toy, column="subtype", subtypes=c("A", "B", "C"))

# Select top genes by median absolute deviation
genes.selected <- rank.var(toy, column="subtype", subtypes=c("A", "B", "C"), genes=genes.filtered, limit=75)

# Subset toy dataset
eset <- toy[genes.selected,]
```

### Structure Constraints
```{r, results='hide'}
# Detect modules
wgcna <- mods.detect(eset, min.size=5, cor.fn="cor", do.plot=FALSE)
```

```{r}
# Module membership
mods.plot(wgcna$dat, wgcna$mods, wgcna$colors, ncol=3, size=2.5)
```

```{r}
# Module extension
mods.extended <- fuzzy.mods(wgcna$dat, wgcna$mods, p=0.75)
mods.extended$grey <- NULL
mods <- sort(sapply(wgcna$mods, length), decreasing=TRUE)
print(mods)
sapply(mods.extended, length)[names(mods)]
```

### Network Estimation

```bash
$ curl -s https://get.nextflow.io | bash
```

--
*Hint*
Once downloaded make the `nextflow` file accessible by your $PATH variable so you do not have to specify the full path to nextflow each time. e.g. `nextflow run` rather than `path/to/nextflow run`
--

**Clone**
*See full documentation for [shine-nf](https://github.com/montilab/shine-nf).*
```bash
$ git clone https://github.com/montilab/shine-nf
```

**Docker**  
```bash
$ docker pull montilab/shine:latest
```

**Hierarchy**  


[![](https://mermaid.ink/img/eyJjb2RlIjoiJSV7aW5pdDogeyd0aGVtZSc6ICdiYXNlJywgJ3NlY3VyaXR5TGV2ZWwnOiAnbG9vc2UnLCAndGhlbWVWYXJpYWJsZXMnOiB7ICdwcmltYXJ5Q29sb3InOiAnIzc4NjhlNjQwJywgJ2VkZ2VMYWJlbEJhY2tncm91bmQnOicjZmZmZmZmJywgJ3RlcnRpYXJ5Q29sb3InOiAnI2Y0ZjRmNCd9fX0lJVxuXG5ncmFwaCBURFxuICAgIFNDe3tDb25zdHJhaW50c319IC0tPiB8TGVhcm58IEFCQyhcIkFCQyA8YnIvPiBuPTMwXCIpXG4gICAgQUJDIC0uLT4gfFByaW9yIExlYXJufCBBQltcIkFCIDxici8-IG49MjBcIl1cbiAgICBBQiAtLi0-IHxQcmlvciBMZWFybnwgQVtcIkEgPGJyLz4gbj0xMFwiXVxuICAgIEFCIC0uLT4gfFByaW9yIExlYXJufCBCW1wiQiA8YnIvPiBuPTEwXCJdXG4gICAgQUJDIC0uLT4gfFByaW9yIExlYXJufCBDW1wiQyA8YnIvPiBuPTEwXCJdXG4gIFxuICAgIHN1YmdyYXBoIFwiIFwiXG4gICAgQVxuICAgIEJcbiAgICBDXG4gICAgZW5kXG4iLCJtZXJtYWlkIjp7InRoZW1lIjoiZGVmYXVsdCJ9LCJ1cGRhdGVFZGl0b3IiOmZhbHNlfQ)](https://mermaid-js.github.io/mermaid-live-editor/#/edit/eyJjb2RlIjoiJSV7aW5pdDogeyd0aGVtZSc6ICdiYXNlJywgJ3NlY3VyaXR5TGV2ZWwnOiAnbG9vc2UnLCAndGhlbWVWYXJpYWJsZXMnOiB7ICdwcmltYXJ5Q29sb3InOiAnIzc4NjhlNjQwJywgJ2VkZ2VMYWJlbEJhY2tncm91bmQnOicjZmZmZmZmJywgJ3RlcnRpYXJ5Q29sb3InOiAnI2Y0ZjRmNCd9fX0lJVxuXG5ncmFwaCBURFxuICAgIFNDe3tDb25zdHJhaW50c319IC0tPiB8TGVhcm58IEFCQyhcIkFCQyA8YnIvPiBuPTMwXCIpXG4gICAgQUJDIC0uLT4gfFByaW9yIExlYXJufCBBQltcIkFCIDxici8-IG49MjBcIl1cbiAgICBBQiAtLi0-IHxQcmlvciBMZWFybnwgQVtcIkEgPGJyLz4gbj0xMFwiXVxuICAgIEFCIC0uLT4gfFByaW9yIExlYXJufCBCW1wiQiA8YnIvPiBuPTEwXCJdXG4gICAgQUJDIC0uLT4gfFByaW9yIExlYXJufCBDW1wiQyA8YnIvPiBuPTEwXCJdXG4gIFxuICAgIHN1YmdyYXBoIFwiIFwiXG4gICAgQVxuICAgIEJcbiAgICBDXG4gICAgZW5kXG4iLCJtZXJtYWlkIjp7InRoZW1lIjoiZGVmYXVsdCJ9LCJ1cGRhdGVFZGl0b3IiOmZhbHNlfQ)

**Define Workflow**
```sh
#!/usr/bin/env nextflow
# wf.nf

workflow ABC {
    main:
      eset = "data/esets/ABC.rds"
      modules = "data/modules.rds"
      SPLIT( eset, modules )
      LEARN( SPLIT.out.flatten() )
      RECONSTRUCT( eset, LEARN.out[0].collect() )
    emit:
      LEARN.out[0]
}
workflow AB {
    take: 
      prior
    main:
      eset = "data/esets/AB.rds"
      LEARN_PRIOR( eset, prior )
      RECONSTRUCT( eset, LEARN_PRIOR.out[0].collect() )
     emit:
     LEARN_PRIOR.out[0]
}
workflow A {
    take: 
      prior
    main:
      eset = "data/esets/A.rds"
      LEARN_PRIOR( eset, prior )
      RECONSTRUCT( eset, LEARN_PRIOR.out[0].collect() )
}
workflow B {
    take: 
      prior
    main:
      eset = "data/esets/B.rds"
      LEARN_PRIOR( eset, prior )
      RECONSTRUCT( eset, LEARN_PRIOR.out[0].collect() )
}
workflow C {
    take: 
      prior
    main:
      eset = "data/esets/C.rds"
      LEARN_PRIOR( eset, prior )
      RECONSTRUCT( eset, LEARN_PRIOR.out[0].collect() )
}
workflow {
    ABC()
    AB(ABC.out)
    A(AB.out)
    B(AB.out)
    C(ABC.out)
}
```

**Run**  
```bash
$ nextflow run wf.nf -with-docker montilab/shine
```

```
N E X T F L O W  ~  version 20.07.1
Launching `wf.nf` [happy_koch] - revision: c2526aec9e
executor >  local (36)
[f4/c715fd] process > ABC:SPLIT          [100%] 1 of 1 ✔
[1b/37cc11] process > ABC:LEARN (6)      [100%] 6 of 6 ✔
[99/23dcba] process > ABC:RECONSTRUCT    [100%] 1 of 1 ✔
[1b/4e3537] process > AB:LEARN_PRIOR (6) [100%] 6 of 6 ✔
[2a/42f4df] process > AB:RECONSTRUCT     [100%] 1 of 1 ✔
[da/17bb2f] process > A:LEARN_PRIOR (6)  [100%] 6 of 6 ✔
[59/d209b2] process > A:RECONSTRUCT      [100%] 1 of 1 ✔
[26/65e7a0] process > B:LEARN_PRIOR (6)  [100%] 6 of 6 ✔
[7e/b65391] process > B:RECONSTRUCT      [100%] 1 of 1 ✔
[58/721e46] process > C:LEARN_PRIOR (6)  [100%] 6 of 6 ✔
[99/517354] process > C:RECONSTRUCT      [100%] 1 of 1 ✔
Completed at: 21-Nov-2020 16:00:29
Duration    : 2m 16s
CPU hours   : 0.1
Succeeded   : 36
```
