---
output: rmarkdown::github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo=FALSE, message=FALSE}
knitr::opts_chunk$set(message=FALSE, comment="#>")
devtools::load_all(".")
```

# shine

[![](https://img.shields.io/badge/platforms-linux%20%7C%20osx%20-2a89a1.svg)]()
[![](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Built With](https://img.shields.io/badge/Built%20With-Nextflow-brightgreen.svg)](https://www.nextflow.io/)
[![](https://img.shields.io/github/last-commit/montilab/shine.svg)](https://github.com/montilab/shine/commits/master)

Bayesian **S**tructure Learning for **Hi**erarchical **Ne**tworks  
*A package to aid in the Bayesian estimation of hierarchical biological regulatory networks via Gaussian graphical modeling*

## Documentation

Please visit https://montilab.github.io/shine/ for comprehensive documentation.

## Requirements

We suggest R 3.6.0 but R (>= 3.5.0) is required to install directly from Github. For workflows, you will need Python (>= 2.7.0) and dependencies for [Nextflow](https://www.nextflow.io/).  

Nextflow can be used on any POSIX compatible system (Linux, OS X, etc) and requires BASH and Java 8 (or higher) to be installed. Alternatively, check out usage with Docker.

## Installation

Install the development version of the package from Github.

```{r, eval=FALSE}
devtools::install_github("montilab/shine")
```

## Basic Overview

```{r, eval=FALSE}
library(shine)
```

### Example Data
```{r}
# Toy expression set object
data(toy)
```

```{r}
dim(toy)
table(toy$subtype)
```

### Variable Selection
```{r}
# Filter out non-varying genes
genes.filtered <- filter.var(toy, column="subtype", subtypes=c("A", "B", "C"))

# Select top genes by median absolute deviation
genes.selected <- select.var(toy, column="subtype", subtypes=c("A", "B", "C"), genes=genes.filtered, limit=150)

# Subset toy dataset
eset <- toy[genes.selected,]
```

### Structural Constraints
```{r, results='hide'}
# Find constraints for structure with zero-order correlations
wgcna <- mods.detect(eset, min.size=5, cores=3, do.plot=FALSE)
metanet <- bdg.metanet(wgcna$eigengenes, cut=0.5, mpl=TRUE, iter=20000, cores=3)

# Set constraints
blanket <- blanket.new(wgcna$genes)
blanket <- blanket.lift(blanket, wgcna$mods, metanet$edges)
```

### Hierarchical Network Workflows
```{r, eval=FALSE}
condition <- "subtype"

# Learn networks as a hierarchy
hierarchy <- "
A_B_C -> A_B
A_B_C -> C
A_B -> A
A_B -> B
"

# Generate workflow
build.workflow(hierarchy,
               condition,
               eset,
               blanket,
               iter=10000,
               cores=3)
```

```bash
./nextflow workflow.nf -c workflow.config -profile local
```

```md
N E X T F L O W  ~  version 19.10.0
Launching `workflow.nf` [mad_northcutt] - revision: e7aff2bf48
-
W O R K F L O W ~ Configuration
===============================
data      : /Users/anthonyfederico/Downloads/nextflow/data/data.rds
output    : /Users/anthonyfederico/Downloads/nextflow
-------------------------------

Hierarchy
A_B_C -> A_B
A_B_C -> C
A_B -> A
A_B -> B

-

executor >  local (5)
[4d/715ff1] process > A_B_C [100%] 1 of 1 ✔
[87/c2d03b] process > C     [100%] 1 of 1 ✔
[6b/544fdd] process > A_B   [100%] 1 of 1 ✔
[96/916564] process > A     [100%] 1 of 1 ✔
[43/c599d4] process > B     [100%] 1 of 1 ✔

Completed at: 24-Dec-2019 23:01:05
Duration    : 1m 45s
CPU hours   : 0.1
Succeeded   : 5
```

### Network Visualization

Check out https://github.com/montilab/netviz to explore your networks.
