---
title: "Network Learning"
vignette: >
  %\VignetteIndexEntry{Network learning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(comment="", cache=FALSE)
devtools::load_all(".")
```

# Simplified Example

This is the last step in learning networks. At this point, you should have one or more datasets to learn networks for, employed a gene selection method, and learned structural constraints for your networks.

```{r, eval=FALSE}
library(shine)
```

```{r}
data(toy)
```

```{r}
dim(toy)
```

```{r, eval=FALSE}
# Filter out non-varying genes
genes.filtered <- keep.var(toy, column="subtype", subtypes=c("A", "B", "C"))

# Select top genes by median absolute deviation
genes.selected <- rank.var(toy, column="subtype", subtypes=c("A", "B", "C"), genes=genes.filtered, limit=150)

# Subset toy dataset
eset <- toy[genes.selected,]

# Find constraints for structure with zero-order correlations
wgcna <- mods.detect(eset, min.size=5, cores=3, do.plot=FALSE)
fuzzy <- fuzzy.mods(wgcna$dat, wgcna$mods, p=0.9)
fuzzy$grey <- NULL

# Set constraints
blanket <- blanket.new(wgcna$genes)
blanket <- blanket.lift(blanket, mods=fuzzy)

# Learn network
bdg <- bdg.estimate(data=t(Biobase::exprs(eset)), 
                    mpl=TRUE,
                    g.prior=blanket, 
                    method="ggm", 
                    iter=5000, 
                    print=100, 
                    cores=3)
```

# Network Workflows

Due to the heavy computational requirements for high-dimensional graphical modeling, we learn network hierarchies in parallel on high performance computing platforms. Because some networks are dependent on others (e.g. A network is used as a prior in learning another) we need a reactive workflow to handle all of the processes.

We have create a separate resources for learning networks called [shine-nf](https://github.com/montilab/shine-nf). This is a collection of Nextflow modules (DSL2) for modeling hierarchical biological regulatory networks. It wraps methods developed in the shine R package in Nextflow modules for building hierarchical workflows.

**Nextflow**  
Workflows are built using [Nextflow](https://www.nextflow.io/). Nextflow can be used on any POSIX compatible system (Linux, OS X, etc) and requires BASH and Java 8 (or higher) to be installed. Download the latest version of Nextflow compatible with DSL2:

```bash
$ curl -s https://get.nextflow.io | bash
```

--
*Hint*
Once downloaded make the `nextflow` file accessible by your $PATH variable so you do not have to specify the full path to nextflow each time. e.g. `nextflow run` rather than `path/to/nextflow run`
--

**Clone Directory**  
```bash
$ git clone https://github.com/montilab/shine-nf
```

**Docker**  
```bash
$ docker pull montilab/shine:latest
```

## Example Hierarchy

```{r}
dim(toy)
table(toy$subtype)
```

```
    ABC
    / \
   AB  \ 
  /  \  \
 A    B  C 
```

### Individual / No Constraints

This most simple workflow includes building networks for groups A, B, and C without any prior information.

```sh
workflow A {
    main:
      eset = "data/esets/A.rds"
      LEARN( eset )
}
workflow B {
    main:
      eset = "data/esets/B.rds"
      LEARN( eset, prior )
}
workflow C {
    main:
      eset = "data/esets/C.rds"
      LEARN( eset, prior )
}
workflow {
    A()
    B()
    C()
}
```

### Individual / Constraints

This most simple workflow includes building networks for groups A, B, and C without any prior information.

```sh
workflow A {
    main:
      eset = "data/esets/A.rds"
      modules = "data/modules.rds"
      SPLIT( eset, modules )
      LEARN( SPLIT.out.flatten() )
      RECONSTRUCT( eset, LEARN.out[0].collect() )
}
workflow B {
    main:
      eset = "data/esets/B.rds"
      modules = "data/modules.rds"
      SPLIT( eset, modules )
      LEARN( SPLIT.out.flatten() )
      RECONSTRUCT( eset, LEARN.out[0].collect() )
}
workflow C {
    main:
      eset = "data/esets/C.rds"
      modules = "data/modules.rds"
      SPLIT( eset, modules )
      LEARN( SPLIT.out.flatten() )
      RECONSTRUCT( eset, LEARN.out[0].collect() )
}
workflow {
    A()
    B()
    C()
}
```

### Hierarchy / Constraints
```sh
workflow ABC {
    main:
      eset = "data/esets/ABC.rds"
      modules = "data/modules.rds"
      SPLIT( eset, modules )
      LEARN( SPLIT.out.flatten() )
      RECONSTRUCT( eset, LEARN.out[0].collect() )
    emit:
      LEARN.out[0]
}
workflow AB {
    take: 
      prior
    main:
      eset = "data/esets/AB.rds"
      LEARN_PRIOR( eset, prior )
      RECONSTRUCT( eset, LEARN_PRIOR.out[0].collect() )
     emit:
     LEARN_PRIOR.out[0]
}
workflow A {
    take: 
      prior
    main:
      eset = "data/esets/A.rds"
      LEARN_PRIOR( eset, prior )
      RECONSTRUCT( eset, LEARN_PRIOR.out[0].collect() )
}
workflow B {
    take: 
      prior
    main:
      eset = "data/esets/B.rds"
      LEARN_PRIOR( eset, prior )
      RECONSTRUCT( eset, LEARN_PRIOR.out[0].collect() )
}
workflow C {
    take: 
      prior
    main:
      eset = "data/esets/C.rds"
      LEARN_PRIOR( eset, prior )
      RECONSTRUCT( eset, LEARN_PRIOR.out[0].collect() )
}
workflow {
    ABC()
    AB(ABC.out)
    A(AB.out)
    B(AB.out)
    C(ABC.out)
}
```

Refer to [shine-nf](https://github.com/montilab/shine-nf) for detailed documentation.
